---
title: "ESM204-Assignment3"
author: "Geoffrey Cook"
date: "5/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Reading in Libraries and Databases}
##############################################################
# Reading in libraries and databases
##############################################################

# Libraries
library(tidyverse)
library(RColorBrewer)
library(stargazer)
library(vcdExtra)

# Data
whales <- read_csv("HW3Data.csv")

```

```{r Question 1}
##############################################################
# Question 1
##############################################################
# Create a linear probability model that predicts a respondent’s probability of voting ‘yes’ on the ballot based on their age, income, NEP score, the risk reduction offered by the program, and the cost of the program to that respondent. Show the model and interpret the regression coefficients
##############################################################

#Data exploring
whale_plot <- ggplot(whales, aes(x = NEP, y = vote)) +
  geom_point(aes(color = income))
whale_plot


#Check classs of categorical variables
class(whales$age) 
class(whales$income) 
class(whales$income) 

#Change class to "factor"
whales$age <- factor(whales$age)
whales$income <- factor(whales$income)

#Change reference levels
whales$age <- fct_relevel(whales$age, "tothirty", "toforty", "tofifty", "tosixty", "oversixty")
whales$income <- fct_relevel(whales$income, "poor", "middle", "rich", "very_rich", "one_percent")

#Check reference levels
levels(whales$age)
levels(whales$income)


# Linear regression for probability of voting 'YES'
prob_lm <- lm(vote ~ age+income+NEP+risk+bid, data = whales)
summary(prob_lm)

#Model results:
#P(yes) =  0.1428764 + 0.0158639*NEP + 0.0007445*risk - 0.0010699*bid - 0.0027386*middle + 0.0047505*rich + 0.0440535*veryrich + 0.0060895*onepercent - 0.0405591*toforty - 0.0104585*tofifty - 0.0366662*tosixty - 0.0204401*oversixty


```



```{r Question 2}
##############################################################
# Question 2
##############################################################
# Based on this regression, what can you say about the value of a single prevented whale death? (Hint: Think about how risk reduction for all whales translates into the number of whale deaths avoided)
##############################################################

#20% reduction --> 5 less whale deaths
#4% reduction --> 1 less whale death

#Q: How does bid change if the risk reduction is 4%?

#Bid function

#vote =  0.1428764 + 0.0158639*NEP + 0.0007445*risk - 0.0010699*bid - 0.0027386*middle + 0.0047505*rich + 0.0440535*veryrich + 0.0060895*onepercent - 0.0405591*toforty - 0.0104585*tofifty - 0.0366662*tosixty - 0.0204401*oversixty

# 0.0010699*bid = 0.1428764 + 0.0158639*NEP + 0.0007445*risk - 0.0027386*middle + 0.0047505*rich + 0.0440535*veryrich + 0.0060895*onepercent - 0.0405591*toforty - 0.0104585*tofifty - 0.0366662*tosixty - 0.0204401*oversixty - vote


bidfxn <- function(NEP, risk, middle, rich, veryrich, onepercent, toforty, tofifty, tosixty, oversixty, vote) {(0.1428764 + 0.0158639*NEP + 0.0007445*risk - 0.0027386*middle + 0.0047505*rich + 0.0440535*veryrich + 0.0060895*onepercent - 0.0405591*toforty - 0.0104585*tofifty - 0.0366662*tosixty - 0.0204401*oversixty - vote) / 0.0010699}

#Averages
mean_risk <- mean(whales$risk) #49
mean_NEP <- mean(whales$NEP) #38.366
mean_vote <- mean(whales$vote) #0.714
summary(whales$age) #Mode is tothirty
summary(whales$income) #Mode is one_percent

#If the level of risk reduction is 20%, what would the bid be?
bidfxn(NEP = 38.366, risk = 20, middle = 0, rich = 0, veryrich = 0, onepercent = 1, toforty = 0, tofifty = 0, tosixty = 0, oversixty = 0, vote = 0.714) # 54.66893

#If the level of risk reduction is 24%, what would the bid be?
bidfxn(NEP = 38.366, risk = 24, middle = 0, rich = 0, veryrich = 0, onepercent = 1, toforty = 0, tofifty = 0, tosixty = 0, oversixty = 0, vote = 0.714) # 57.45237

#So then what is the value of 1 whale? 
value1 <- 57.45237 - 54.66893
# $2.78344

```




Question 3: 

Pick three arbitrary respondents. For each of those respondents, estimate their WTP for a VSR
program offering 60% risk reduction.

```{r}
# Linear regression for WTP
wtp_lm <- lm(bid ~ age+income+NEP+risk+vote, data = whales)
summary(wtp_lm)
# wtp = 69.46411 - 0.04160*NEP - 0.040145*risk - 5.02885*vote - 1.04449*middle - 0.54404*rich - 2.19112*veryrich - 2.35561*onepercent - 3.45011*toforty - 2.55066*tofifty - 6.60334*tosixty - 1.83242*oversixty

wtpfxn <- function(NEP, risk, vote, middle, rich, veryrich, onepercent, toforty, tofifty, tosixty, oversixty) {69.46411 - 0.04160*NEP - 0.040145*risk - 5.02885*vote - 1.04449*middle - 0.54404*rich - 2.19112*veryrich - 2.35561*onepercent - 3.45011*toforty - 2.55066*tofifty - 6.60334*tosixty - 1.83242*oversixty}


#(48, very_rich, toforty)
wtpfxn(NEP = 48, risk = 60, middle = 0, rich = 0, veryrich = 1, onepercent = 0, toforty = 1, tofifty = 0, tosixty = 0, oversixty = 0, vote = 0.7) 
# 55.89719

#(45,one_percent, tosixty)
wtpfxn(NEP = 45, risk = 60, middle = 0, rich = 0, veryrich = 0, onepercent = 1, toforty = 0, tofifty = 0, tosixty = 1, oversixty = 0, vote = 0.7) 
# 52.70427

#(22, poor, tothirty)
wtpfxn(NEP = 22, risk = 60, middle = 0, rich = 0, veryrich = 0, onepercent = 0, toforty = 0, tofifty = 0, tosixty = 0, oversixty = 0, vote = 0.7) 
# 62.62002

```
Question 4.

Repeat this process and estimate the mean WTP for a VSR program offering 60% risk
reduction among Santa Barbara County households.

```{r}

#Use average values of dataset to find mean WTP

mean_wtp <- wtpfxn(NEP = mean_NEP, risk = 60, middle = 0, rich = 0, veryrich = 0, onepercent = 1, toforty = 0, tofifty = 0, tosixty = 0, oversixty = 0, vote = mean_vote) 
# Mean WTP = 59.51318
```

Question 5-6.
If there are 150,000 households in Santa Barbara County, estimate the total benefit of a VSR
program to these households. Based only on these benefits, do the benefits of the VSR program outweigh the costs?
```{r}
mean_wtp*150000
total_benefit <- 8926976
total_costs <- 7000000
net_total <- total_benefit - total_costs
net_total # 1,926,976
#Benefits > costs
```

Question 7.
Suppose the price of a carbon credit is $Z per ton, at what price Z* will the shipping industry
decide to adopt the VSR for purely self-interested reasons? (Note: if ships voluntarily reduce
speed, then the $7 million cost (noted above) of implementing the VSR would be avoided)

```{r}
#Ships will voluntarily adopt the policy of the cost of speed reduction < costs of carbon credit

1000 < 20 * Z
Z > 50

#Z must be at least $50 per ton of carbon
```


Question 8
Now suppose the carbon credit price is exactly Z*, so all ships voluntarily reduce speed to
achieve the 60% risk reduction. Approximately how many whales would be saved by this? What
would be the social value of allowing ships to enter the carbon trading market?

```{r}
# 20% risk reduction -> 5 whales saved
# 60% risk reduction-> 15 whales saved


```

#Social value
Costs/benefits of voluntary program:
Costs = $7 million to implement 
Benefits = $8,926,976

Costs/benefits of cap and trade program:
Costs = $1000 per
Benefits = $8,926,976


?????????????????????



